---
title: "Module8"
author: "Natalie Robinson"
date: "September 25, 2017"
output: html_document
---
# Homework Problem:
Every Saturday, at the same time, a primatologist goes and sits in the forest in the morning and listens for titi monkey calls, counting the number of calls they hear in a 2 hour window from 5am to 7am. Based on previous knowledge, she believes that the mean number calls she will hear in that time is exactly 15. Let X represent the appropriate Poisson random variable of the number of calls heard in each monitoring session.

What is the probability that she will hear more than 8 calls during any given session? *****
```{r}
x <- 8
l <- 15
probset1 <- 1- ppois(x, l)
probset1
```

What is the probability that she will hear no calls in a session?
```{r}
x <- 0
l <- 15
probset2 <- dpois(x, l)
probset2
```

What is the probability that she will hear exactly 3 calls in a session?
```{r}
x <- 3
l <- 15
probset3 <- dpois(x, l)
probset3
```

Plot the relevant Poisson mass function over the values in range 0 ≤ X ≤ 30.
```{r}
x <- 0:30
l = 15
graphprobset <- dpois(x = x, lambda = l)
barplot(graphprobset, names.arg = x, space = 0, xlab = "x", ylab = "Pr(X=x)", main = "Probability Mass Function")
```

Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions). 
```{r}
n <- 104
l <- 15
results <- rpois(n = n, lambda = l)
results
```

Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. 
```{r}
hist(results, xlim = c(0,30), probability = TRUE, main = "Simulated Results", 
    xlab = "number of calls heard", ylab = "probability")
```

How does your histogram compare to the shape of the probability mass function you plotted above?
* The shape of the histogram is relatively similar to the shape of the probability mass function, as they both are somewhat bell shaped with a peak around x = 15 (the mean). Both graphs have lower probability values towards the upper and lower limits of the x-axis. 

